---
title: "group-summary"
author: "Meredith Miller, Carter Burtlake, Naomi Lubkin, Declan Taylor"
date: "01/11/2021"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}
library(here)
library(popbio)
library(vegan)
library(labdsv)
library(tidyverse)
```

## Introduction

How does intertidal height affect biodiversity?

variables needed: height at each transect

**Hypothesis**: differences in shore elevation and wave exposure affect the total biodiversity present due to differences in emersion time and wave splash.

## Data Analysis

"Your analyses and displayed results must relate to the hypotheses you made about the effect of habitat structure on intertidal community structure at the outset of the project. The challenge here is to decide on how best to present your results- as is the case with all primary research. Begin by exploring the data, making some preliminary graphs, having discussions with your teammates, and discussing challenges with your instructors. You will need to decide on the best approach for displaying the main results. Your group should create 2-3 clear graphs that visually display your results, and one summary table of the results of accompanying statistical analysis."

### Exploring our data

#### Importing the data
```{r}
biodiversity <- read_csv(here("./Data/biodiversity.csv"))

percent_cover <- read_csv(here("./Data/percent-cover.csv")) 

quadrat_data <- read_csv(here("./Data/quadrat-data.csv"))
```

#### Modifying the data

Please note that tidal heights were calculated using the Canadian Hydrographic Service's 2021 Tide and Current Tables (Vol. 5)

```{r, echo = FALSE}
# Create a data.frame with the tidal height values for each quadrant
tidal_height <- quadrat_data %>%
  group_by(vertical_transect, quadrat) %>%
  summarize("quadrat_height_cm" = quadrat_height)

# Biodiversity is calculated using the diversity() function from the vegan
# package. The `biodiversity` dataframe needs to be reformatted to wide data
# to allow the "vegan" package to work.

# Setting up the `biodiversity` data.frame to the "3-column" format required by
# the matrify() function from the "labdsv" package.
matrify_biodiv <- biodiversity %>%
  drop_na(n) %>%
  group_by(vertical_transect, quadrat) %>%
  
  # Generating a unique ID that contains both the `vertial_transect` and 
  # `quadrat` information to be read by the matrify() function, which requires 
  # site to be contained within 1 column.
  mutate(unique_ID = as.numeric(
                        paste0(vertical_transect, ".", quadrat))) %>%
  select(unique_ID, species, n)

# Dropping the group by columns with irrelevant information
matrify_biodiv <- data.frame(
  matrify_biodiv[,3:5])

# Running the matrify function to create a data.frame that can be read by the
# "vegan" diversity() function.
vegan_biodiv <- matrify(matrify_biodiv)

# Save the Shannon Index data as a data.frame
ShannonIndex <- data.frame(
  diversity(vegan_biodiv, index = "shannon"))

# Modifications to the Shannon Index data
ShannonIndex <- ShannonIndex %>%
  # Turning the row names back into a column
  cbind(UniqueID = rownames(ShannonIndex)) %>%
  # Separating UniqueID back into distinct transect and quadrat numbers.
  separate(UniqueID,
           c("vertical_transect", "quadrat")) %>%
  # Adding quadrant height and wave exposure to the dataframe
  cbind("quadrat_height_cm" = tidal_height$quadrat_height_cm) %>%
  cbind("exposure" = quadrat_data$side) %>%
  #Fixing the column names.
  rename("ShannonIndex" = diversity.vegan_biodiv..index....shannon..)

```

#### Visualising the data
```{r}
# Create a plot
DiversityHeight_Plot <-
  ggplot(data = ShannonIndex,
         aes(x = quadrat_height_cm,
             y = ShannonIndex)) +
  geom_point(aes(colour = exposure)) +
  geom_smooth(colour = "black",
              size = 0.7,
              method = lm, se = FALSE) +
  labs(x = "Quadrat Height (cm)",
       y = "Shannon Index",
       colour = "Site Exposure") +
  theme_classic()

DiversityHight_Plot
```






Code we're probs not gonna use:
>>>>>>> 569c66a5936ae0ad2573724a0b796f229229f3f7
```{r}
# Cleaning the percent cover CSV file
percent_cover <- percent_cover %>%
  mutate(per_sand = per_sand/100,
         per_pebble = per_pebble/100,
         per_cobble = per_cobble/100,
         per_boulder = per_boulder/100,
         per_bedrock = per_bedrock/100,
         per_mussel = per_mussel/100,
         per_barnacle = per_barnacle/100,
         per_upright = per_upright/100,
         per_crust = per_crust/100)

# Adding a column o calculate species richness
tidal_height <- quadrat_data %>%
  group_by(vertical_transect, quadrat) %>%
  summarize("quadrat_height_cm" = quadrat_height)

quad_richness <- biodiversity %>%
  drop_na(n) %>%
  group_by(vertical_transect, quadrat) %>%
  summarize(n_spp = length(unique(species))) %>%
  # Match the quadrat to a tidalhight value
  cbind("quadrat_height_cm" = tidal_height$quadrat_height_cm) %>%
  cbind("exposure" = quadrat_data$side)

spp_evenness <- biodiversity %>%
  drop_na(n) %>%
  group_by(vertical_transect, quadrat) %>%
  summarize(n_indv = length(unique(species))) %>%
  cbind("quadrat_height_cm" = tidal_height$quadrat_height_cm) %>%
  cbind("exposure" = quadrat_data$side)
```

